\textsc{\wobble: a data-driven \changed{analysis technique for time-series stellar spectra}}

### 2019-08-29 paper/paper.tex
math tweaks

First, we assume that the wavelength calibration and spectral extraction of the instrument are perfect: that is, we begin at the stage of having 1D extracted spectra and corresponding wavelength grids in hand and we do not model any corrections to the wavelength solution. 
\changed{Similarly, we assume that the line spread function of the instrument remains perfectly constant from one exposure to the next. This assumption is needed because it allows us to extract instrumentally-broadened template spectra rather than explicitly modeling and solving for instrumental broadening effects. However, as we discuss in Section \ref{s:data-changes}, even within the limits of this assumption extracting already-broadened templates is not strictly correct. We do this for simplicity only and leave the general case to future work.}

@@

\changed{For each data column \mvec{y_n}, our model prediction \mvec{f_n} can be treated as the sum of stellar and telluric contributions:
\begin{equation}
\mvec{f_n} = \mvec{f_{\star,n}} + \mvec{f_{t,n}} + \textrm{noise}.
\end{equation}}

\changed{The stellar spectrum contribution is: 
\begin{equation}
\mvec{f_{\star,n}} = \mmatrix{P}(\mvec{\xi_n}, \mvec{\xi_{\star}}, v_{obs,n}) \cdot \mvec{\mu_{\star}},
\label{eqn:star}
\end{equation}
where \mvec{\mu_{\star}} is a spectral template of log-fluxes and 
\mvec{\xi_{\star}} is the corresponding vector of template log-wavelengths. 
The exact values of \mvec{\xi_{\star}} can be chosen somewhat arbitrarily under the conditions that the grid has uniform spacing $\Delta\xi_{\star}$, covers the entire wavelength range of the data, and is over-sampled with regards to the observed spectrum's wavelength grid.
\mmatrix{P} is a linear operator whose function is to apply a Doppler shift by observed velocity $v_{obs,n}$ and
interpolate \mvec{\mu_{\star}} from the \mvec{\xi_{\star}} template grid to the \mvec{\xi_n} data grid.
Each entry of the \mmatrix{P} matrix can be defined by a sum of weighted indicator functions (where an indicator function is denoted here as $\mathbb{1}(x)$ and is defined to have value 1 when condition $x$ is fulfilled and zero otherwise):
\begin{equation}
\begin{split}
 P_{i,j} =  &\ \Big(\frac{\xi_{n,i} - \xi'_{\star,j}}{\Delta\xi_{\star}}\Big) \mathbb{1}\big(0 < \frac{\xi_{n,i} - \xi'_{\star,j}}{\Delta\xi_{\star}} < 1\big) \\
  &+ \Big(1 - \frac{\xi_{n,i} - \xi'_{\star,j}}{\Delta\xi_{\star}}\Big) \mathbb{1}\big({-1} < \frac{\xi_{n,i} - \xi'_{\star,j}}{\Delta\xi_{\star}} < 0\big), \\
 \end{split}
\end{equation}
where \mvec{\xi'_{\star}} is the Doppler-shifted template grid:
\begin{equation}
 \mvec{\xi'_{\star}}(v) = \mvec{\xi_{\star}} + D(v) \equiv \mvec{\xi_{\star}} + \frac{1}{2} \ln \left(\frac{1 - v/c}{1 + v/c}\right).
\end{equation}}
Because \mmatrix{P} is quite sparse, in practice we do not instantiate the full matrix when performing calculations. 
\changed{At this point, it is also useful to note that \mmatrix{P} could, in principle, encode an instrumental line spread function (\LSF). 
We return to this point in Section \ref{s:future}.}

@@

We evaluate a log-likelihood for each epoch
\changed{
\begin{equation}
\ln \mathcal{L}_n = -\frac{1}{2} (\mvec{y_{n}} - \mvec{f_{n}})^T \mvec{C_{n}}^{-1} (\mvec{y_{n}} - \mvec{f_{n}}),
\label{eqn:lnlike}
\end{equation}}
comparing our model \changed{\mmatrix{f}} to the data \mmatrix{y}, with \mmatrix{C} representing the covariance matrix of uncertainties for the data.

@@

L1 normalization adds a term to the log-likelihood that takes the form:
\begin{equation}
\Lone{\lambda}{p} \equiv -\lambda \sum_{i} | p_{i} | ,
\end{equation}
where \mvec{p} is the vector of parameters to be normalized (in this case \mvec{\mu_{\star}}, \mvec{\mu_{t}}, or \mmatrix{W_{t}}) and $\lambda$ is the regularization amplitude.

Similarly, L2 normalization adds a term of the form:
\begin{equation}
\Ltwo{\lambda}{p} \equiv -\lambda \sum_{i} p_{i}^2 .
\end{equation}

>>>>

First, we assume that the wavelength calibration and spectral extraction of the instrument are perfect: that is, we begin at the stage of having 1D extracted spectra and corresponding wavelength grids in hand and we do not model any corrections to the wavelength solution. 
\changed{Similarly, we assume that the line spread function of the instrument remains perfectly constant from one exposure to the next. This assumption is needed because it allows us to extract constant instrumentally-broadened template spectra rather than explicitly modeling and solving for time-variable broadening effects. However, as we discuss in Section \ref{s:data-changes}, even within the limits of this assumption extracting already-broadened templates is not strictly correct. We do this for simplicity only and leave the general case to future work.}

@@

\changed{For each data column \mvec{y_n}, our model prediction \mvec{f_n} can be treated as the sum of stellar and telluric contributions at time $n$:
\begin{equation}
\mvec{f_n} = \mvec{f_{\star,n}} + \mvec{f_{t,n}} + \textrm{noise}.
\end{equation}}

\changed{The stellar spectrum contribution is: 
\begin{equation}
\mvec{f_{\star,n}} = \mmatrix{P}(\mvec{\xi_n}, \mvec{\xi_{\star}}, v_{obs,n}) \cdot \mvec{\mu_{\star}},
\label{eqn:star}
\end{equation}
where \mvec{\mu_{\star}} is a spectral template of log-fluxes and 
\mvec{\xi_{\star}} is the corresponding vector of template log-wavelengths. 
The exact values of \mvec{\xi_{\star}} can be chosen somewhat arbitrarily under the conditions that the grid has uniform spacing $\Delta\xi_{\star}$, covers the entire wavelength range of the data, and is over-sampled with regards to the observed spectrum's wavelength grid.
\mmatrix{P} is a linear operator whose function is to apply a Doppler shift by observed velocity $v_{obs,n}$ and
interpolate \mvec{\mu_{\star}} from the \mvec{\xi_{\star}} template grid to the \mvec{\xi_n} data grid.
Each entry of the \mmatrix{P} matrix can be defined by a sum of weighted indicator functions (where an indicator function is denoted here as $\mathbb{1}(x)$ and is defined to have value 1 when condition $x$ is fulfilled and zero otherwise):
\begin{equation}
\begin{split}
 P_{i,j} =  &\ \bigg(\frac{\xi_{n,i} - \xi'_{\star,j}}{\Delta\xi_{\star}}\bigg) \cdot \mathbb{1}\bigg(0 \leq \frac{\xi_{n,i} - \xi'_{\star,j}}{\Delta\xi_{\star}} < 1\bigg) \\
  &+ \bigg(1 - \frac{\xi_{n,i} - \xi'_{\star,j}}{\Delta\xi_{\star}}\bigg) \cdot \mathbb{1}\bigg({-1} < \frac{\xi_{n,i} - \xi'_{\star,j}}{\Delta\xi_{\star}} \leq 0\bigg), \\
 \end{split}
\end{equation}
where \mvec{\xi'_{\star}} is the Doppler-shifted template grid:
\begin{equation}
 \mvec{\xi'_{\star}}(v) = \mvec{\xi_{\star}} + D(v) \equiv \mvec{\xi_{\star}} + \frac{1}{2} \ln \left(\frac{1 - v/c}{1 + v/c}\right).
\end{equation}}
Because \mmatrix{P} is quite sparse, in practice we do not instantiate the full matrix when performing calculations. 
\changed{At this point, it is also useful to note that \mmatrix{P} could, in principle, encode an instrumental line spread function (\LSF). 
We return to this point in Section \ref{s:future}.}

@@

\changed{The contribution of each data epoch to the net log-likelihood may be evaluated as
\begin{equation}
\ln \mathcal{L}_n = -\frac{1}{2} (\mvec{y_{n}} - \mvec{f_{n}})^T \mvec{C_{n}}^{-1} (\mvec{y_{n}} - \mvec{f_{n}}),
\label{eqn:lnlike}
\end{equation}
}with \mmatrix{C_n} representing the covariance matrix of uncertainties on the data.

@@

L1 normalization adds a term to the log-likelihood that takes the form:
\begin{equation}
\Lone{\lambda}{p} \equiv -\lambda \sum_{i} | p_{i} | ,
\end{equation}
where \mvec{p} is the vector of parameters to be normalized (in this case \mvec{\mu_{\star}}, \mvec{\mu_{t}}, or \mmatrix{W_{t}}) and $\lambda$ is the regularization amplitude.
Similarly, L2 normalization adds a term of the form:
\begin{equation}
\Ltwo{\lambda}{p} \equiv -\lambda \sum_{i} p_{i}^2 .
\end{equation}
